### 2022-05-12 make/makefile.tex
Initial commit

# This is a makefile template to be copied & modified. Copy it to
# "Makefile" in your directory and do "chmod u+w Makefile".
# -- Andrew Myers

DOC = paper
#
# change this line to contain the name of the document without extension

TARGETS = $(DOC).pdf
#
# TARGETS could be $(DOC).dvi instead, if there are no figures

TEXS = $(DOC).tex *.tex *.sty
#
# Add any included TeX files to TEXS

FIGS =
#
# List of postscript figures to be included

PAPERSROOT=..
MAKEDIR=$(PAPERSROOT)/make

include $(MAKEDIR)/commondefs
#
# Change MAKEDIR to . if you have commondefs installed in the current
# directory.

default: $(TARGETS)

BIBFILE = $(DOC).bib
# BIBFILE = $(PAPERSROOT)/bibtex/master.bib
#
# Ignore this if you're not using bibtex. Change the $(DOC) part if your bib
# file has a different name from your document

$(DOC).dvi: $(TEXS) $(FIGS) $(DOC).stamp
$(DOC).pdf: $(TEXS) $(FIGS) $(DOC).stamp
#
# To use bibtex, add these to dependency list:
# $(DOC).bbl $(BIBFILE)
# 
# To use makeindex, add this to dependency list:
# $(DOC).ind
#
# To use glossaries, add this to dependency list:
# $(DOC).glo

$(DOC).bbl: $(BIBFILE) $(DOC).stamp
$(DOC).glo $(DOC).ind: $(DOC).stamp
$(DOC).ps: $(DOC).dvi $(FIGS)
#
# These are standard dependencies. Shouldn't need to modify these.

VPATH = $(PAPERSROOT)/tex-macros

include $(COMMONRULES)