### 2022-05-12 tex-macros/aplcomments-test.tex

\documentclass[12pt]{article}
\usepackage{times,graphicx}
\IfFileExists{.comments}{
\typeout{** Showing comments **}
\usepackage{hyperref}
\usepackage{aplcomments}
}{
\usepackage[disabled]{aplcomments}
}
%\usepackage[inline]{aplcomments}
\textwidth 7in
\advance\oddsidemargin -1.25in
\usepackage{aplcomments}
\newcommenter{Andrew}{0.0,0.5,1.0}
\begin{document}

This song was written for the Cornell Computer Science Department
holiday party in December 2009. It was performed by Andrew Myers,
with Dexter Kozen accompanying\Andrew{awesomely}.

\Andrew{Lyrics by Andrew Myers.}

This is another paragraph.
\Andrew{and another comment}

And here is a third paragraph, which should be indented even if
comments are being used.

\Andrew{A comment before a section heading might get eaten}

\section{Lyrics follow this point}

\Andrew{A comment after a section heading should be okay}

{
\obeylines
\setlength\parskip{0.7em}
\Andrew{The beginning}
A new student came by just the other day
to start some research right away
But there were papers to review and grants to raise
and he took courses while I was away
And he gave talks 'fore I knew it, and good ones too
He'd say, ``I'm gonna be like you, yeah"
"You know I'm gonna be like you"
\hrule
Refrain:
\fbox{
\begin{minipage}{\linewidth}
The committee meetings and the blind reviews,
Writing homeworks and exams, and grading them too.
``When can we meet, Professor?"
``Don't know when, but we'll do some research then"
``You know we'll do some nice work then"
\end{minipage}
}
My student started third year just the other day
He said, ``My first paper's going well now that I've got an RA"
``Can you help me write the intro today?"
I said, ``Not right now, I have exams to grade"
And he walked away but his smile never dimmed
He said, ``I'm gonna be like him, yeah"
``You know I'm gonna be like him"
\hrule
(Refrain)
\hrule
Well, he finished his thesis just the other day
Such a brilliant researcher I just had to say,
``Hey, I'm proud of you, want to talk for a while?"
He shook his head and then said with a smile
``What I really need now is a very strong letter"
``See you later, can you write it soon, please?
See you later, can you write it soon?"
\hrule
And as I sent off the email it occurred to me
He'd ended up just like me.
Yeah, my student was just like me.
}

\begin{figure}

\begin{center}
\Andrew{comment in a figure}
\fbox{Figure}
\end{center}
\caption{This is a caption\Andrew{and a comment in a caption}}
\label{fig}
\end{figure}

\ref{fig}
final paragraph

\end{document}