### 2022-10-12 stlc-isorec-prod.tex

\begin{figure}
\lstset{
    basicstyle=\fontsize{8}{8.5}\ttfamily,
% numbers=left,
}

\begin{minipage}{\textwidth}
\begin{multicols}{2}

\definecolor{codecomment-color}{HTML}{DD033F}

\begin{lstlisting}
Family STLCIsorec extends STLC.[@\tanc{STLCIsorec-start}@]
[@\hfill\codecomment{STLC extended with iso-recursive types}@]
FInductive ty : Set +=
| ty_var : id -> ty | ty_rec : id -> ty -> ty.

FInductive tm : Set +=
| tm_fold : tm -> tm | tm_unfold : tm -> tm.

FRecursion tysubst on ty[@\hfill\codecomment{Type-level substitution}@]
motive \lambda(_ : ty), id -> ty -> ty.
Case ty_unit  := [@\dadada@].   Case ty_arrow := [@\dadada@].
Case ty_var   := [@\dadada@].   Case ty_rec   := [@\dadada@].
End tysubst.

FInductive hasty : env -> tm -> ty -> Prop +=
| ht_fold : \forall G t \alpha T,
    hasty G t (tysubst T \alpha (ty_rec \alpha T)) ->
    hasty G (tm_fold t) (ty_rec \alpha T)
| ht_unfold : [@\dadada@].

[@\dadada@][@\hfill\codecomment{Other adjustments}@]
End STLCIsorec.[@\tanc{STLCIsorec-end}@]
\end{lstlisting}

\makeline[0pt]{STLCIsorec-start}{STLCIsorec-end}[codecomment-color!50]

\vspace{-12pt}

\definecolor{codecomment-color}{HTML}{008B8B}

\begin{lstlisting}
Family STLCFixIsorec extends STLC[@\tanc{STLCIsorecfix-start}@]
using STLCFix, STLCIsorec.
[@\hfill\ \ \codecomment{STLC extended with fixpoints and iso-recursive types}@]
End STLCFixIsorec.[@\tanc{STLCIsorecfix-end}@]
\end{lstlisting}

\makeline[0pt]{STLCIsorecfix-start}{STLCIsorecfix-end}[codecomment-color!50]

\columnbreak

\definecolor{codecomment-color}{HTML}{0DA3FF}

\begin{lstlisting}
Family STLCProd extends STLC.[@\tanc{STLCProd-start}@]
[@\hfill\codecomment{STLC extended with products}@]
FInductive ty : Set +=
| ty_prod : ty -> ty -> ty.

FInductive tm : Set +=
| tm_pair : tm -> tm -> tm
| tm_fst : tm -> tm | tm_snd : tm -> tm.

[@\dadada@][@\hfill\codecomment{Other adjustments}@]
End STLCProd.[@\tanc{STLCProd-end}@]
\end{lstlisting}

\makeline[.5\textwidth+9pt]{STLCProd-start}{STLCProd-end}[codecomment-color!50]

\vspace{-10pt}

\definecolor{codecomment-color}{HTML}{E8A20C}

\begin{lstlisting}
Family STLCProdIsorec extends STLC[@\tanc{STLCIsorecProd-start}@]
using STLCProd, STLCIsorec.
[@\hfill\ \ \codecomment{STLC extended with products and iso-recursive types}@]
FRecursion tysubst on ty
motive \lambda(_ : ty), id -> ty -> ty.
Case ty_prod := [@\dadada@].[@\hfill\codecomment{Substitution on product types}@]
End tysubst.

End STLCProdIsorec.[@\tanc{STLCIsorecProd-end}@]
\end{lstlisting}

\makeline[.5\textwidth+9pt]{STLCIsorecProd-start}{STLCIsorecProd-end}[codecomment-color!50]

\vspace{-10pt}

\definecolor{codecomment-color}{HTML}{708090}

\begin{lstlisting}
Family STLCFixProdIsorec extends STLC[@\tanc{STLCIsorecFixProd-start}@]
using STLCFix, STLCProdIsorec.[@\hfill\codecommentstart{STLC extended}@]
[@\hfill\codecommentend{with fixpoints, products, and iso-recursive types}@]
End STLCFixProdIsorec.[@\tanc{STLCIsorecFixProd-end}@]
\end{lstlisting}

\makeline[.5\textwidth+9pt]{STLCIsorecFixProd-start}{STLCIsorecFixProd-end}[codecomment-color!50]

\end{multicols}
\end{minipage}

\vspace{-10pt}

\caption{Composing extensions of STLC.}
\label{fig:stlc-isorec-prod}
\end{figure}