### 2022-10-16 stlc-compiled.tex
simplify

\begin{figure}
\vspace{-5pt}
\begin{minipage}{\textwidth}
\begin{multicols}{2}

\lstset{
  basicstyle=\fontsize{8}{8.2}\ttfamily,
% numbers=left,
}

\colorlet{codecomment-color}{dkgreen}

\newcommand{\innersep}{\vspace{6pt}}
\newcommand{\outersep}{\vspace{5pt}}
\newcommand{\outdent}{\!\!\!}

\begin{lstlisting}
[@\outdent@][@\codecomment{Code emitted upon definition of \texttt{tm} in family \texttt{STLC}}@][@\tanc{c-STLC-tm-start}@]
[@\outdent@]Module Type STLC\_tm\_Ctx.
[@\outdent@]End STLC\_tm\_Ctx.[@\innersep@]
[@\outdent@]Module Type STLC\_tm (self : STLC\_tm\_Ctx).
Axiom tm : Set.
Axiom tm_unit : tm.[@\hfill@]Axiom tm_var : id -> tm.
Axiom tm_abs : id -> tm -> tm.
Axiom tm_app : tm -> tm -> tm.
Axiom tm_prect_STLC : [@\dadada@].
Axiom [@\textls[-30]{tm\_unit\_eq\_STLC}@][@\;:\;@][@\dadada@]. Axiom [@\textls[-30]{tm\_abs\_eq\_STLC}@][@\;:\;@][@\dadada@].
Axiom [@\textls[-30]{tm\_abs\_eq\_STLC}@][@\;:\;@][@\dadada@].  Axiom [@\textls[-30]{tm\_app\_eq\_STLC}@][@\;:\;@][@\dadada@].
[@\outdent@]End STLC\_tm.[@\tanc{c-STLC-tm-end}@][@\outersep@]
[@\outdent@][@\codecomment{Code emitted upon definition of \texttt{subst} in family \texttt{STLC}}@][@\tanc{c-STLC-subst-start}@]
[@\outdent@]Module Type STLC\_subst\_Ctx.
Include STLC\_tm\_Ctx.  Include STLC\_tm.
[@\outdent@]End STLC\_subst\_Ctx.[@\innersep@]
[@\outdent@]Module STLC\_subst\_Cases (self : STLC\_subst\_Ctx).
Def subst\_tm_unit :=
  \lambda (x : id) (t : self.tm), self.tm_unit.
Def subst\_tm_var := [@\dadada@].
Def subst\_tm_abs := [@\dadada@].  Def subst\_tm_app := [@\dadada@].
[@\outdent@]End STLC\_subst\_Cases.[@\innersep@]
[@\outdent@]Module Type STLC\_subst (self : STLC\_subst\_Ctx).
Axiom subst : self.tm[@\,@]->[@\,@]id[@\,@]->[@\,@]self.tm[@\,@]->[@\,@]self.tm.
Axiom subst_tm_unit_eq : \forall x t,
  self.subst (self.tm_unit) x t = self.tm_unit.
Axiom subst_tm_var_eq : [@\dadada@].
Axiom subst_tm_abs_eq : [@\dadada@].
Axiom subst_tm_app_eq : [@\dadada@].
[@\outdent@]End STLC\_subst.[@\tanc{c-STLC-subst-end}@][@\outersep@]
[@\outdent@][@\dadada@][@\hfill\codecomment{Code emitted upon definition of \texttt{ty} in family \texttt{STLC}}@][@\tanc{c-STLC-ty-start}@][@\tanc{c-STLC-ty-end}@][@\outersep@]
[@\outdent@][@\codecomment{Code emitted upon definition of \texttt{env} in family \texttt{STLC}}@][@\tanc{c-STLC-env-start}@]
[@\outdent@]Module Type STLC\_env\_Ctx.
Include STLC\_ty\_Ctx.  Include STLC\_ty.
[@\outdent@]End STLC\_env\_Ctx.[@\innersep@]
[@\outdent@]Module STLC\_env (self : STLC\_env\_Ctx).
Def env : Type := id -> option self.ty.
[@\outdent@]End STLC\_env.[@\tanc{c-STLC-env-end}@][@\outersep@]
[@\outdent@][@\dadada@][@\hfill\codecomment{Code emitted for other fields defined in family \texttt{STLC}}@][@\tanc{c-STLC-rest}@][@\outersep@]
[@\outdent@][@\codecomment{Code emitted upon completion of family \texttt{STLC}}@][@\tanc{c-STLC-start}@]
[@\outdent@]Module STLC.[@\innersep@]
Inductive tm\_ : Set :=[@\tanc{c-STLC-d-tm-start}@]
| tm_unit\_[@\,@]:[@\,@]tm\_ | tm_var\_[@\,@]:[@\,@]id -> tm\_
| tm_abs\_[@\,@]:[@\,@]id->tm\_->tm\_ | tm_app\_[@\,@]:[@\,@]tm\_->tm\_->tm\_.
Def tm := tm\_.[@\hfill\codecomment{Instantiate \texttt{tm} \& its constructors}@]
Def tm_unit := tm_unit\_. Def tm_var := tm_var\_.
Def tm_abs  := tm_abs\_.  Def tm_app := tm_app\_.
[@\hfill\codecomment{Instantiate \texttt{tm} partial recursor \& its equalities}@]
Def tm_prect_STLC P := tm\__rect (\lambda[@\,@]t,[@\,@]option[@\,@](P[@\,@]t)).
Fact[@\;@][@\textls[-20]{tm\_unit\_eq\_STLC}@][@\,@]:[@\,@][@\dadada@].[@\,@][@\textls[-40]{{\color{dkblue}reflexivity}.}@][@\,@]Qed.[@\hfill@]Fact [@\dadada@][@\tanc{c-STLC-d-tm-end}@][@\innersep@]
Include STLC\_subst\_Cases.[@\tanc{c-STLC-d-subst-start}@]
Def subst[@\,@] :=[@\hfill\codecomment{Instantiate \texttt{subst} \& its equalities}@]
  tm\__rect _ subst\_tm_unit subst\_tm_var
             subst\_tm_abs  subst\_tm_app.
Fact[@\;@][@\textls[-40]{subst\_tm\_unit\_eq}@][@\,@]:[@\,@][@\dadada@].[@\,@][@\textls[-40]{{\color{dkblue}reflexivity}.}@][@\,@]Qed.[@\hfill@]Fact [@\dadada@][@\tanc{c-STLC-d-subst-end}@][@\tanc{c-STLC-firstcol-end}\columnbreak@][@\innersep@]
[@\dadada@][@\hfill\codecomment{Instantiate \texttt{ty}, its constructors, partial recusor \& eq.}@][@\tanc{c-STLC-d-ty}@][@\tanc{c-STLC-sndcol-start}@][@\innersep@]
Include STLC\_env.[@\hfill\codecomment{Include \texttt{env}}@][@\tanc{c-STLC-d-env}@][@\innersep@]
[@\dadada@][@\hfill\codecomment{Include/Instantiate other fields of family \texttt{STLC}}@][@\tanc{c-STLC-d-rest}@][@\innersep@]
[@\outdent@]End STLC.[@\tanc{c-STLC-end}@]
\end{lstlisting}

\makeline[-1pt]{c-STLC-subst-start}{c-STLC-subst-end}[codecomment-color!55][1.8pt]
\makeline[-1pt]{c-STLC-tm-start}{c-STLC-tm-end}[codecomment-color!55][1.8pt]
\makeline[-1pt]{c-STLC-ty-start}{c-STLC-ty-end}[codecomment-color!55][1.8pt]
\makeline[-1pt]{c-STLC-env-start}{c-STLC-env-end}[codecomment-color!55][1.8pt]
\makeline[-1pt]{c-STLC-rest}{c-STLC-rest}[codecomment-color!55][1.8pt]
\makeline[-1pt]{c-STLC-start}{c-STLC-firstcol-end}[codecomment-color!55][1.8pt]
\makeline[4pt]{c-STLC-d-tm-start}{c-STLC-d-tm-end}[codecomment-color!35][1.0pt]
\makeline[4pt]{c-STLC-d-subst-start}{c-STLC-firstcol-end}[codecomment-color!35][1.0pt]
\makeline[.5\textwidth+8pt]{c-STLC-sndcol-start}{c-STLC-end}[codecomment-color!55][1.8pt]
\makeline[.5\textwidth+12pt]{c-STLC-d-ty}{c-STLC-d-ty}[codecomment-color!35][1.0pt]
\makeline[.5\textwidth+12pt]{c-STLC-d-env}{c-STLC-d-env}[codecomment-color!35][1.0pt]
\makeline[.5\textwidth+12pt]{c-STLC-d-rest}{c-STLC-d-rest}[codecomment-color!35][1.0pt]

\vspace{-12pt}

\colorlet{codecomment-color}{orange}

\begin{lstlisting}
[@\outdent@][@\codecomment{Code emitted upon definition of \texttt{tm} in family \texttt{STLCFix}}@][@\tanc{c-STLCFix-tm-start}@]
[@\outdent@]Module Type STLCFix\_tm\_Ctx.
[@\outdent@]End STLCFix\_tm\_Ctx.[@\innersep@]
[@\outdent@]Module Type STLCFix\_tm (self : STLCFix\_tm\_Ctx).
Include STLC\_tm(self).
Axiom tm_fix : id -> tm -> tm.
Axiom tm_fix_eq_STLC : \forall[@\,@][@\dadada@],[@\,@]tm_prect_STLC [@\dadada@][@\;@]=[@\;@]None.
Axiom tm_prect_STLCFix : [@\dadada@].
Axiom tm_fix_eq_STLCFix : [@\dadada@].
[@\outdent@]End STLCFix\_tm.[@\tanc{c-STLCFix-tm-end}@][@\outersep@]
[@\outdent@][@\codecomment{Code emitted upon definition of \texttt{subst} in family \texttt{STLCFix}}@][@\tanc{c-STLCFix-subst-start}@]
[@\outdent@]Module Type STLCFix\_subst\_Ctx.
Include STLCFix\_tm\_Ctx. Include STLCFix\_tm.
[@\outdent@]End STLCFix\_subst\_Ctx.[@\innersep@]
[@\outdent@]Module STLCFix\_subst\_Cases (self[@\,@]:[@\,@]STLCFix\_subst\_Ctx).
Include STLC\_subst\_Cases(self).
Def subst\_tm_fix := [@\dadada@].
[@\outdent@]End STLCFix\_subst\_Cases.[@\innersep@]
[@\outdent@]Module Type STLCFix\_subst (self[@\,@]:[@\,@]STLCFix\_subst\_Ctx).
Include STLC\_subst(self).
Axiom subst_tm_fix_eq : [@\dadada@].
[@\outdent@]End STLCFix\_subst.[@\tanc{c-STLCFix-subst-end}@][@\outersep@]
[@\outdent@][@\dadada@] [@\codecomment{Code emitted for other fields defined in family \texttt{STLCFix}}@][@\tanc{c-STLCFix-rest}@][@\outersep@]
[@\outdent\codecomment{Code emitted upon completion of family \texttt{STLCFix}}@][@\tanc{c-STLCFix-start}@]
[@\outdent@]Module STLCFix.[@\innersep@]
Inductive tm\_ : Set :=[@\tanc{c-STLCFix-d-tm-start}@]
| tm_unit\_[@\,@]:[@\,@]tm\_ | tm_var\_[@\,@]:[@\,@]id -> tm\_
| tm_abs\_[@\,@]:[@\,@]id->tm\_->tm\_ | tm_app\_[@\,@]:[@\,@]tm\_->tm\_->tm\_
| tm_fix\_[@\,@]:[@\,@]id->tm\_->tm\_.
Def tm := tm\_.[@\hfill\codecomment{Instantiate \texttt{tm} \& its constructors}@]
Def tm_unit := tm_unit\_. Def tm_var := tm_var\_.
Def tm_abs  := tm_abs\_.  Def tm_app := tm_app\_.
Def tm_fix  := tm_fix\_.
[@\dadada@][@\hfill\codecomment{Instantiate \texttt{tm} partial recursors \& their equalities}@][@\tanc{c-STLCFix-d-tm-end}@][@\innersep@]
Include STLCFix\_subst\_Cases. [@\tanc{c-STLCFix-d-subst-start}@]
Def subst :=[@\hfill\codecomment{Instantiate \texttt{subst} \& its equalities}@]
  tm\__rect _ subst\_tm_unit subst\_tm_var
             subst\_tm_abs  subst\_tm_app subst\_tm_fix.
Fact[@\;@][@\textls[-40]{subst\_tm\_unit\_eq}@][@\,@]:[@\,@][@\dadada@].[@\,@][@\textls[-40]{{\color{dkblue}reflexivity}.}@][@\,@]Qed.[@\hfill@]Fact [@\dadada@][@\tanc{c-STLCFix-d-subst-end}@][@\innersep@]
[@\dadada@][@\hfill\codecomment{\textls[-10]{Instantiate \texttt{ty}, its constructors, partial recursor \& eq.}}@][@\tanc{c-STLCFix-d-ty}@][@\innersep@]
Include STLC\_env.[@\hfill\codecomment{Include \texttt{env}}@][@\tanc{c-STLCFix-d-env}@][@\innersep@]
[@\dadada@][@\hfill\codecomment{Include/Instantiate other fields of family \texttt{STLCFix}}@][@\tanc{c-STLCFix-d-rest}@][@\innersep@]
[@\outdent@]End STLCFix.[@\tanc{c-STLCFix-end}@]
\end{lstlisting}

\makeline[.5\textwidth+8pt]{c-STLCFix-tm-start}{c-STLCFix-tm-end}[codecomment-color!55][1.8pt]
\makeline[.5\textwidth+8pt]{c-STLCFix-subst-start}{c-STLCFix-subst-end}[codecomment-color!55][1.8pt]
\makeline[.5\textwidth+8pt]{c-STLCFix-rest}{c-STLCFix-rest}[codecomment-color!55][1.8pt]
\makeline[.5\textwidth+8pt]{c-STLCFix-start}{c-STLCFix-end}[codecomment-color!55][1.8pt]
\makeline[.5\textwidth+13pt]{c-STLCFix-d-tm-start}{c-STLCFix-d-tm-end}[codecomment-color!35][1.0pt]
\makeline[.5\textwidth+13pt]{c-STLCFix-d-subst-start}{c-STLCFix-d-subst-end}[codecomment-color!35][1.0pt]
\makeline[.5\textwidth+13pt]{c-STLCFix-d-ty}{c-STLCFix-d-ty}[codecomment-color!35][1.0pt]
\makeline[.5\textwidth+13pt]{c-STLCFix-d-env}{c-STLCFix-d-env}[codecomment-color!35][1.0pt]
\makeline[.5\textwidth+13pt]{c-STLCFix-d-rest}{c-STLCFix-d-rest}[codecomment-color!35][1.0pt]

\vspace{-12pt}

\definecolor{codecomment-color}{HTML}{000080}

\begin{lstlisting}
[@\outdent@][@\codecomment{Code emitted upon command \texttt{\textls[-35]{Check STLCFix.typesafe}}}@][@\tanc{c-Check-STLCFix-typesafe-start}@]
[@\outdent@]Check STLCFix.typesafe.[@\tanc{c-Check-STLCFix-typesafe-end}@]
\end{lstlisting}

\makeline[.5\textwidth+8pt]{c-Check-STLCFix-typesafe-start}{c-Check-STLCFix-typesafe-end}[codecomment-color!50][1.8pt]

\end{multicols}
\end{minipage}

%\vspace{-10pt}

\caption{Compilation of the mechanization of the STLCs in \cref{fig:stlc-mechanized} to parameterized Coq modules.}
\label{fig:stlc-compiled}
\end{figure}