### 2022-10-23 metatheory.tex
FIx language: distinguish judgement and the derivation of the judgement.

% Motivation of using QIIT
We start with the formulation of the syntax in the declarative style, which is based on the
formalization of a predicative \textit{Martin-Lof Type Theory} (MLTT)~\cite{martin1982constructive}
given by~\citet{coquand2018canonicity}. We will mainly focus on the new facility that we introduced.

The reason we choose this version of predicative MLTT is mainly because it is simpler on the formulation and metatheoretical reasoning, easy to extend and is enough as a background for setting up the formalization of the key feature for our language, like overriding and inheritance about the families. Compared to a more complete Coq's formal foundation like \textit{Predicative Calculus of Cumulative Inductive Constructions}~\cite{timany2018cumulative}, our formulation lacks cumulativity (though our formulation still has multi-level universe) and impredicative universe (the \mintinline{Coq}{Prop : Type}). \YZ{
    What's the relation to the CIC underlying Coq, given that the implementation is a Coq plugin?
}\EDJreply{Add the comparison between this version of predicative MLTT and PCUIC. Please Check. }
% Thus we will omit a great deal of details about MLTT here

Though phrased in Latex, we are following~\citet{altkap2016} to use as
our meta-logic an extensional type theoretical framework (with Agda-like syntax), 
together with \textit{Quotient Inductive Inductive Type} (QIIT).

\textit{Quotient Inductive Inductive Type}~(QIIT) is an extension of \textit{inductive-inductive type} (IIT)~\cite{nordvall2010inductive}. IIT can be roughly considered as several inductive type mutually defined, in a more expressible way than mutual-inductive type. QIIT is allowing to have \textit{quotient}, in other words, equality results as part of the constructors. For example, in \\ \mintinline{Agda}{data N2: Set where O : N2 | S : N2 → N2 | B : S (S x) ≡ x }, the "B" here is a valid (equality) constructor, and "N2" is a set of two elements in our extensional setting.

Our formulation can be seen as following the practice ``type theory should eat itself''\cite{dybjer1995internal, chapman2009type}. We use as the formulation the intrinsically-typed debruijn syntax in this type-theoretical QIIT instead of a more conventional and untyped representation, because (1) we hope our formulation can be precise on the mathematical object we are using (2) it is easier to validate a pen-and-paper proof now because proof checking then is simply type checking in type theory (3) when using QIIT to formulate syntax, the induction principle used in reasoning about syntax is itself justified (4) we hope our formulation can be closer to a real mechanized one so that the reader can have an idea about how mechanization can be carried out (5) We can still use named-variables for intuition and example illustration (6) using QIIT as syntax encoding can avoid the reader from the concept of \textit{presupposition}~\cite{gratzer-sterling-birkedal-2019}---these \textit{presupposition} can be understood as \textit{implicit parameters} of the declaration of the typing rules in QIIT.

% In our syntax, each judgement will actually be represented by a QIIT type, which are
% mutually recursively defined together. Quotient is used to represent
% judgemental(definitional) equality, thanks to which we can have a
% concise representation. This way, for all four kinds of judgement we are
% using, there are always equality among these data. Doing so, we can omit
% some obvious equality judgement. Most of the reasoning can be formulated
% using the \textit{algebra} of QIIT (mapping out function from this
% quotient data-type).

% We also highlight some special feature in this style of formulation of dependent type theory: 1. this modern formulation of type theory does't have operational semantic but only equality 
% %We will recover canonicity afterwards, which implies the computational ability of our theory. 
% ;
% 2. instead of using meta-level substitution, we use explicit substitution and de Bruijn indices in our intrinsic formulation. This is also called substitution calculus in the literature and is favoured due to its clear categorical semantic; 
% 3. we are dealing with quotient data, so we are actually manipulating equivalence classes of terms.

% We will write out the syntax of the whole type theory for completeness.

\subsection{Syntax in Declarative Style}
% \YZ{Rewrite syntax part into name-variable style instead of currently debruijn indices, and do that for all the models we introduced}

\input{syntax}

\subsection{Implementation Formalized as Syntactic Translation}

\input{syn-translate.tex}

\subsection{Standard Model for Consistency}

\input{standardmodel.tex}

\subsection{Proof Relevant Logical Relation for Canonicity}
\input{canonicitymodel.tex}