### 2022-10-31 stlc-linkage-typing.tex
figure

\begin{figure}
\small
\renewcommand*{\arraystretch}{1.15}
\newcolumntype{L}{>{$}l<{$}}
\newcolumntype{C}{>{$}c<{$}}
\newcolumntype{R}{>{$}r<{$}}
\begin{tabular}{@{}lL@{\ }@{}}
\lsti!Family STLC.!
&
\lsig_0 \coloneqq \LSigEmp
\quad
\lkg_0 \coloneqq \LkgEmp
\quad
\goodTerm{\cdot}{\lkg_1}{\TyLkg{\lsig_1}}
\\[4pt]
\multirow{2}{*}{\lsti!FInductive tm :=!}
&
\lsig_1 \coloneqq \LSigAdd{\lsig_0}{\seal_1}{T_1}
\quad
\lkg_1 \coloneqq \LkgAdd{\lkg_0}{t_1}
\quad
\goodTerm{\cdot}{\lkg_2}{\TyLkg{\lsig_2}}
\quad
\goodTerm{\TyTkg{\lsig_0}}{\seal_1}{\sub{A_1}{\SubstWeak{1}}}
\\
&
\goodTerm{\mathit{self} : \big[\big]}{\wcode{\wsig_{\texttt{tm}}}}{\TyS{\wcode{\wsig_{\texttt{tm}}}}}
\\[4pt]
\multirow{2}{*}{\lsti!| tm_unit : tm |!\ \dadada}
&
\lsig_2 \coloneqq \LSigAdd{\lsig_1}{\seal_2}{T_2}
\quad
\lkg_2 \coloneqq \LkgAdd{\lkg_1}{t_2}
\quad
\goodTerm{\cdot}{\lkg_2}{\TyLkg{\lsig_2}}
\quad
\goodTerm{\TyTkg{\lsig_1}}{\seal_2}{\sub{A_2}{\SubstWeak{1}}}
\\
&
\goodTerm{
    \mathit{self} : \big[\texttt{tm} : \TyS{\wcode{\wsig_{\texttt{tm}}}}\big]
}{
    \Wsup{\wsig_{\texttt{tm}}}{\top}{\bot}
}{
    \El{\mathit{self.}\texttt{tm}}
}
\\[4pt]
\lsti!FRecursion subst!\ \dadada
\\
\multirow{2}{*}{\lsti!Case tm_unit :=!\ \dadada}
&
\lsig_6 \coloneqq \LSigAdd{\lsig_5}{\seal_6}{T_6}
\quad
\lkg_6 \coloneqq \LkgAdd{\lkg_5}{t_6}
\quad
\goodTerm{\cdot}{\lkg_6}{\TyLkg{\lsig_6}}
\quad
\goodTerm{\TyTkg{\lsig_5}}{\seal_6}{\sub{A_6}{\SubstWeak{1}}}
\\
&
\goodTerm{
    \mathit{self} : \big[\texttt{tm} : \cU,\ \texttt{tm\_unit} : \El{\texttt{tm}},\ \dadada\big]
}{
%   \lambda\_.\,\lambda\_.\,\lambda x.\,\lambda t.\,\mathit{self}.\texttt{tm\_unit}
    \dadada
}{
    \CaseSig{\top}{\bot}{\El{\mathit{self}.\texttt{tm}}\to\mathit{id}\to\El{\mathit{self}.\texttt{tm}}}
}
\\[4pt]
\lsti!Case!\ \dadada
\\
\multirow{2}{*}{\lsti!End subst.!}
&
\lsig_{10} \coloneqq \LSigAdd{\lsig_9}{\seal_{10}}{T_{10}}
\quad
\lkg_{10} \coloneqq \LkgAdd{\lkg_9}{t_{10}}
\quad
\goodTerm{\cdot}{\lkg_{10}}{\TyLkg{\lsig_{10}}}
\quad
\goodTerm{\TyTkg{\lsig_9}}{\seal_{10}}{\sub{A_{10}}{\SubstWeak{1}}}
\\
&
\goodTerm{
    \mathit{self} : \left[{%
    \begin{minipage}{.35\textwidth}%
        \vspace{-3.5pt}
    \begin{gather*}
    \begin{array}{@{}l@{}}
        \texttt{tm} : \TyS{\wcode{\wsig_{\texttt{tm}}}},\ 
        \texttt{tm\_unit} : \El{\texttt{tm}},\ 
        \dadada
        \\
        \texttt{subst\_tm\_unit} : \CaseSig{\top}{\bot}{\dadada},\ 
        \dadada
    \end{array}
    \end{gather*}
    \end{minipage}
    }\right]
}{
    \lambda t.\,\Wrec{\wsig_{\texttt{tm}}}{\dadada}{t}
}{
    \begin{minipage}{.20\textwidth}%
        \vspace{-3.5pt}\scriptsize
    \begin{gather*}
    \begin{array}{@{}l@{}}
        \El{\mathit{self}.\texttt{tm}} \to \El{\mathit{self}.\texttt{tm}}
        \\
        \to \mathit{id} \to \El{\mathit{self}.\texttt{tm}}
    \end{array}
    \end{gather*}
    \end{minipage}
}
\\
\dadada
\\[4pt]
\lsti!End STLC.!
&
\goodTerm{\cdot}{\Tkg{\lkg_{20}}}{\TyTkg{\lsig_{20}}}
\end{tabular}

\caption{Typing of each field}
\label{fig:stlc-linkage-typing}
\end{figure}