### 2022-11-01 stlc-linkage-typing.tex
coloring

\begin{figure}
\small
\renewcommand*{\arraystretch}{1.15}
\newcolumntype{L}{>{$}l<{$}}
\newcolumntype{C}{>{$}c<{$}}
\newcolumntype{R}{>{$}r<{$}}
%
\definecolor{A-color}{HTML}{E7EFC5}
\definecolor{T-color}{HTML}{FFF1D0}
\definecolor{t-color}{HTML}{FAE0E4}
\colorlet{Bkg-color}{black!8}
%
\begin{tabular}{@{}l@{\ \ \ }L@{}}
\cellcolor{Bkg-color}
\lsti!Family STLC.!
&
\lsig_0 \coloneqq \LSigEmp
\quad
\lkg_0 \coloneqq \LkgEmp
\quad
\goodTerm{\cdot}{\lkg_1}{\TyLkg{\lsig_1}}
\\
\\[-8pt]
\cellcolor{Bkg-color}
&
\lsig_1 \coloneqq \LSigAdd{\lsig_0}{\seal_1}{
    \tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T1-a}(-0.02,-0.10)(-0.02,0.28)
    T_1
    \tikzmarkend{T1-a}
}
\quad
\lkg_1 \coloneqq \LkgAdd{\lkg_0}{
    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t1-a}(-0.02,-0.10)(-0.02,0.28)
    t_1
    \tikzmarkend{t1-a}
}
\quad
\goodTerm{\cdot}{\lkg_2}{\TyLkg{\lsig_2}}
\quad
\goodTerm{\TyTkg{\lsig_0}}{\seal_1}{\sub{
    \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A1-a}(-0.02,-0.10)(-0.05,0.28)
    A_1
    \tikzmarkend{A1-a}
}{\SubstWeak{1}}}
\\
\cellcolor{Bkg-color}
\multirow{-2}{*}{%
\lsti!FInductive!\ \ 
    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t1-c}(0.01,-0.10)(-0.02,0.28)%
    \lsti!tm!%
    \tikzmarkend{t1-c}%
\ \lsti!:=!}
&
\goodTerm{
    \mathit{self} : 
    \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A1-b}(0.00,-0.15)(0.00,0.30)%
    \big[\big]
    \tikzmarkend{A1-b}%
}{
    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t1-b}(-0.02,-0.10)(-0.02,0.28)
    \wcode{\wsig_{\texttt{tm}}}
    \tikzmarkend{t1-b}
}{
    \tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T1-b}(-0.02,-0.10)(-0.02,0.28)
    \TyS{\wcode{\wsig_{\texttt{tm}}}}
    \tikzmarkend{T1-b}
}
\\
\\[-8pt]
\cellcolor{Bkg-color}
&
\lsig_2 \coloneqq \LSigAdd{\lsig_1}{\seal_2}{
    \tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T2-a}(-0.02,-0.10)(-0.02,0.28)
    T_2
    \tikzmarkend{T2-a}
}
\quad
\lkg_2 \coloneqq \LkgAdd{\lkg_1}{
    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t2-a}(-0.02,-0.10)(-0.02,0.28)
    t_2
    \tikzmarkend{t2-a}
}
\quad
\goodTerm{\cdot}{\lkg_2}{\TyLkg{\lsig_2}}
\quad
\goodTerm{\TyTkg{\lsig_1}}{\seal_2}{\sub{
    \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A2-a}(-0.02,-0.10)(-0.05,0.28)
    A_2
    \tikzmarkend{A2-a}
}{\SubstWeak{1}}}
\\
\cellcolor{Bkg-color}
\multirow{-2}{*}{%
    \lsti!|!\ %
    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t2-c}(0.01,-0.10)(-0.02,0.28)%
    \lsti!tm_unit!%
    \tikzmarkend{t2-c}%
}
&
\goodTerm{
    \mathit{self} :
    \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A2-b}(0.00,-0.15)(0.00,0.30)%
    \big[\texttt{tm} : \TyS{\wcode{\wsig_{\texttt{tm}}}}\big]
    \tikzmarkend{A2-b}%
}{
    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t2-b}(-0.02,-0.10)(-0.02,0.28)
    \Wsup{\wsig_{\texttt{tm}}}{\top}{\bot}
    \tikzmarkend{t2-b}
}{
    \tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T2-b}(-0.02,-0.10)(-0.02,0.28)
    \El{\mathit{self.}\texttt{tm}}
    \tikzmarkend{T2-b}
}
\\
\\[-8pt]
\lsti!| !\dadada\ \ \lsti!| !\dadada\ \ \lsti!| !\dadada\lsti!.!
&
\dadada
\\[4pt]
\lsti!FRecursion subst!\ \dadada\lsti!.!
&
\textit{\color{codecomment-color}%
    $\wcode{\wsig_{\texttt{tm}}}$ is abstracted into $\cU$ in the typing
    context of the case handler below:%
}
\\
\cellcolor{Bkg-color}
&
\lsig_6 \coloneqq \LSigAdd{\lsig_5}{\seal_6}{
    \tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T6-a}(-0.02,-0.10)(-0.02,0.28)
    T_6
    \tikzmarkend{T6-a}
}
\quad
\lkg_6 \coloneqq \LkgAdd{\lkg_5}{
    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t6-a}(-0.02,-0.10)(-0.02,0.28)
    t_6
    \tikzmarkend{t6-a}
}
\quad
\goodTerm{\cdot}{\lkg_6}{\TyLkg{\lsig_6}}
\quad
\goodTerm{\TyTkg{\lsig_5}}{\seal_6}{\sub{
    \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A6-a}(-0.02,-0.10)(-0.05,0.28)
    A_6
    \tikzmarkend{A6-a}
}{\SubstWeak{1}}}
\\
\cellcolor{Bkg-color}
\multirow{-2}{*}{%
    \lsti!Case tm_unit :=!\ %
    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t6-c}(0.01,-0.10)(-0.02,0.28)%
    \dadada%
    \tikzmarkend{t6-c}%
    \lsti!.!%
}
&
\goodTerm{
    \mathit{self} :
    \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A6-b}(0.00,-0.15)(0.00,0.30)%
    \big[\texttt{tm} : \cU,\ \texttt{tm\_unit} : \El{\texttt{tm}},\ \dadada\big]
    \tikzmarkend{A6-b}%
}{
    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t6-b}(-0.02,-0.10)(-0.02,0.28)
%   \lambda\_.\,\lambda\_.\,\lambda x.\,\lambda t.\,\mathit{self}.\texttt{tm\_unit}
    \dadada
    \tikzmarkend{t6-b}
}{
    \tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T6-b}(-0.02,-0.10)(-0.02,0.28)
    \CaseSig{\top}{\bot}{
        \fontsize{8.5}{10.2}\selectfont
        \text{$\El{\mathit{self}.\texttt{tm}}\to\mathit{id}\to\El{\mathit{self}.\texttt{tm}}$}
    }
    \tikzmarkend{T6-b}
}
\\
\\[-8pt]
\lsti!Case!\ \dadada\ \ 
\lsti!Case!\ \dadada\ \ 
\lsti!Case!\ \dadada%
&
\dadada
\\
\cellcolor{Bkg-color}
&
\lsig_{10} \coloneqq \LSigAdd{\lsig_9}{\seal_{10}}{
    \tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T10-a}(-0.02,-0.10)(-0.02,0.28)
    T_{10}
    \tikzmarkend{T10-a}
}
\quad
\lkg_{10} \coloneqq \LkgAdd{\lkg_9}{
    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t10-a}(-0.02,-0.10)(-0.02,0.28)
    t_{10}
    \tikzmarkend{t10-a}
}
\quad
\goodTerm{\cdot}{\lkg_{10}}{\TyLkg{\lsig_{10}}}
\quad
\goodTerm{\TyTkg{\lsig_9}}{\seal_{10}}{\sub{
    \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A10-a}(-0.02,-0.10)(-0.05,0.28)%
    A_{10}
    \tikzmarkend{A10-a}
}{\SubstWeak{1}}}
\\
\cellcolor{Bkg-color}
\multirow{-2}{*}{%
    \lsti!End!\ \ %
    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t10-c}(0.01,-0.10)(-0.02,0.28)%
    \lsti!subst!%
    \tikzmarkend{t10-c}%
    \lsti!.!%
}
&
\goodTerm{
    \mathit{self} :
    \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A10-b}(0.00,-0.32)(0.00,0.47)%
    \!\left[{%
    \begin{minipage}{.31\textwidth}%
    \vspace{-7.5pt}%
    \fontsize{8.0}{9.2}\selectfont%
    \begin{gather*}
    \begin{array}{@{}l@{}}
        \texttt{tm} : \TyS{\wcode{\wsig_{\texttt{tm}}}},\ 
        \texttt{tm\_unit} : \El{\texttt{tm}},\ 
        \dadada,
        \smallskip\\
        \texttt{subst\_tm\_unit} : \CaseSig{\top}{\bot}{\dadada},\ 
        \dadada
    \end{array}
    \end{gather*}
    \end{minipage}
    }\right]\!
    \tikzmarkend{A10-b}%
}{
    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t10-b}(-0.02,-0.10)(-0.02,0.28)
    \lambda t.\,\Wrec{\wsig_{\texttt{tm}}}{\dadada}{t}
    \tikzmarkend{t10-b}
}{
    \begin{minipage}{.20\textwidth}%
    \vspace{-7.5pt}%
    \fontsize{8.0}{9.2}\selectfont%
    \begin{gather*}
    \tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T10-b}(-0.02,-0.26)(-0.02,0.40)
    \begin{array}{@{}l@{}}
        \El{\mathit{self}.\texttt{tm}} \to \El{\mathit{self}.\texttt{tm}}
        \\
        \to \mathit{id} \to \El{\mathit{self}.\texttt{tm}}
    \end{array}
    \tikzmarkend{T10-b}
    \end{gather*}
    \end{minipage}
}
\\
\dadada
&
\dadada
\\
\cellcolor{Bkg-color}
\lsti!End !%
\tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t-STLC-c}(-0.02,-0.10)(-0.02,0.28)%
\lsti!STLC!%
\tikzmarkend{t-STLC-c}%
\lsti!.!
&
\goodTerm{\cdot}{
    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t-STLC-b}(-0.02,-0.10)(-0.02,0.28)%
    \Tkg{\lkg_{20}}
    \tikzmarkend{t-STLC-b}
}{\TyTkg{\lsig_{20}}}
\end{tabular}

\caption{\TT encoding of the \texttt{STLC} family from \cref{fig:stlc-mechanized}.}
\label{fig:stlc-linkage-typing}
\end{figure}