### 2022-11-09 cc-condensed.tex
More Comments.

\noindentparagraph{Consistency.}
One of the most fundamental properties of a dependent type theory is consistency.
\begin{theorem}[Consistency]
\label{thm:consistency}
  The typing judgment $\goodTerm{\cdot}{t}{\bot}$ is not derivable for any term~$t$.
\end{theorem}

\noindent
Consistency says that the type $\bot$ is not inhabited.
Thus, it is safe to use the type theory for logical reasoning, as not
every proposition is trivially provable.
\cref{thm:consistency} is a consequence of a meta\-circular interpretation of all
the well-formedness rules, which, in particular, interprets $\bot$ in
\TT as an uninhabited type in the meta\-language.
The construction of this interpretation follows the standard model of
\citet{altkap2016}, extending it to handle linkages.
Linkages~$\LkgAdd{\lkg}{t}$ can be interpreted into non-dependent products,
as rule \ruleref{l/add} indicates.
%
We refer the reader to a supplemental file (in Agda syntax) for the complete proof.

\noindentparagraph{Canonicity.}

A second property we prove is canonicity.

\begin{theorem}[Canonicity]
\label{thm:canonicity}
  If $\goodTerm{\cdot}{t}{\cB}$, then either $\goodTerm{\cdot}{t \equiv \true}{\cB}$ or $\goodTerm{\cdot}{t \equiv \false}{\cB}$.
\end{theorem}

\noindent
This canonicity theorem says that every closed term of the ground
type~$\cB$ is convertible to one of the canonical forms $\true$ and~$\false$.
When canonicity is proved in a constructive meta\-logic, its proof
amounts to a normalization function for closed terms of the ground type\EDJ{Suggesting use "boolean type". Because natural number type can be also considered ground type in other context (but this theorem on boolean cannot ensure canonicity for nat). We also does specify what is ground type. }.
So canonicity serves to justify the computational nature of the type theory.

We prove \cref{thm:canonicity} by constructing a logical-relations model
for the well-formedness rules, following prior approaches~\cite{coquand2018canonicity,sterling2019algebraic,kaposi2019gluing}.
In particular, a closed, well formed type $\goodType{\cdot}{T}{}$ is interpreted
into a logical predicate on closed terms: the predicate includes all
closed, ``reducible''\EDJ{I will add a it is my bad to introduce "reducible" without defining it first. But my intension is to give a feeling of the dynamic operational semantic for the reader (of course OS is not here in our setting)} terms of type~$T$.
The proof is available as a supplemental file in Agda syntax.