### 2023-03-21 stlc-linkage-typing-named.tex
tweaks to avoid margin overrun

\begin{figure}
\small
\renewcommand*{\arraystretch}{1.15}
\setlength\tabcolsep{0pt}
\newcolumntype{L}{>{$}l<{$}}
\newcolumntype{C}{>{$}c<{$}}
\newcolumntype{R}{>{$}r<{$}}
%
\colorlet{Bkg-color}{black!8}
\colorlet{Bkg-color-2}{white}
%
\begin{tabular}{@{}l@{\ \ \ }L@{\quad}L@{\quad}L@{\quad}L@{}}
\cellcolor{Bkg-color}
\lsti!Family STLC.!
&
\lsig_0 \coloneqq \LSigEmp
&
\lkg_0 \coloneqq \LkgEmp
&
\goodTerm{\cdot}{\lkg_0}{\TyLkg{\lsig_0}}
&
\\
\\[-8pt]
\cellcolor{Bkg-color}
&
\lsig_1 \coloneqq \LSigAdd{\lsig_0}{\seal_1}{
    \tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T1-a}(-0.02,-0.10)(-0.02,0.28)
    T_1
    \tikzmarkend{T1-a}
}
&
\lkg_1 \coloneqq \LkgAdd{\lkg_0}{
    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t1-a}(-0.02,-0.10)(-0.02,0.28)
    t_1
    \tikzmarkend{t1-a}
}
&
\goodTerm{\cdot}{\lkg_1}{\TyLkg{\lsig_1}}
&
\goodTerm{x : \TyTkg{\lsig_0}}{\seal_1}{{
    \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A1-a}(-0.02,-0.10)(-0.05,0.28)
    A_1
    \tikzmarkend{A1-a}
}{}}
\\
\cellcolor{Bkg-color}
\multirow{-2}{*}{%
\lsti!FInductive!\ \ 
    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t1-c}(0.03,-0.10)(-0.02,0.28)%
    \lsti!tm!%
    \tikzmarkend{t1-c}%
\ \;\lsti!:=!}
&
\multicolumn{4}{@{}l@{}}{
\goodTerm{
    \mathit{self} : 
    \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A1-b}(0.00,-0.15)(0.00,0.30)%
    \big[\big]
    \tikzmarkend{A1-b}%
}{
    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t1-b}(-0.02,-0.10)(-0.02,0.28)
    \wcode{\wsig_{\texttt{tm}}}
    \tikzmarkend{t1-b}
}{
    \tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T1-b}(-0.02,-0.10)(-0.02,0.28)
    \TyS{\wcode{\wsig_{\texttt{tm}}}}
    \tikzmarkend{T1-b}
} 
}
\\
\\[-8pt]
\cellcolor{Bkg-color}
&
\lsig_2 \coloneqq \LSigAdd{\lsig_1}{\seal_2}{
    \tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T2-a}(-0.02,-0.10)(-0.02,0.28)
    T_2
    \tikzmarkend{T2-a}
}
&
\lkg_2 \coloneqq \LkgAdd{\lkg_1}{
    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t2-a}(-0.02,-0.10)(-0.02,0.28)
    t_2
    \tikzmarkend{t2-a}
}
&
\goodTerm{\cdot}{\lkg_2}{\TyLkg{\lsig_2}}
&
\goodTerm{x : \TyTkg{\lsig_1}}{\seal_2}{{
    \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A2-a}(-0.02,-0.10)(-0.05,0.28)
    A_2
    \tikzmarkend{A2-a}
}{}}
\\
\cellcolor{Bkg-color}
\multirow{-2}{*}{%
    \lsti!|!\ %
    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t2-c}(0.01,-0.10)(-0.02,0.28)%
    \lsti!tm_unit!%
    \tikzmarkend{t2-c}%
    \ \ \lsti!: tm!
}
&
\multicolumn{4}{@{}l@{}}{
\goodTerm{
    \mathit{self} :
    \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A2-b}(0.00,-0.15)(0.00,0.30)%
    \big[\texttt{tm} : \TyS{\wcode{\wsig_{\texttt{tm}}}}\big]
    \tikzmarkend{A2-b}%
}{
    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t2-b}(-0.02,-0.10)(-0.02,0.28)
    \Wsup{}{\wsig_{\texttt{tm}}}{\top}{\bot}
    \tikzmarkend{t2-b}
}{
    \tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T2-b}(-0.02,-0.10)(-0.02,0.28)
    \El{\mathit{self.}\texttt{tm}}
    \tikzmarkend{T2-b}
}
}
\\
\cellcolor{Bkg-color-2}
\lsti!| !\dadada\ \ \lsti!| !\dadada\ \ \lsti!| !\dadada\lsti!.!
&
\dadada
\\[4pt]
\cellcolor{Bkg-color-2}
\lsti!FRecursion subst!\ \;\dadada\lsti!.!
&
\multicolumn{4}{@{}l@{}}{
\textit{\color{codecomment-color}%
    $\wcode{\wsig_{\texttt{tm}}}$ is abstracted into $\cU$ in the typing
    context (i.e., $A_6$) of the case handler below:%
}
}
\\
\cellcolor{Bkg-color}
&
\lsig_6 \coloneqq \LSigAdd{\lsig_5}{\seal_6}{
    \tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T6-a}(-0.02,-0.10)(-0.02,0.28)
    T_6
    \tikzmarkend{T6-a}
}
&
\lkg_6 \coloneqq \LkgAdd{\lkg_5}{
    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t6-a}(-0.02,-0.10)(-0.02,0.28)
    t_6
    \tikzmarkend{t6-a}
}
&
\goodTerm{\cdot}{\lkg_6}{\TyLkg{\lsig_6}}
&
\goodTerm{x : \TyTkg{\lsig_5}}{\seal_6}{{
    \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A6-a}(-0.02,-0.10)(-0.05,0.28)
    A_6
    \tikzmarkend{A6-a}
}}
\\
\cellcolor{Bkg-color}
\multirow{-2}{*}{%
    \lsti!Case tm_unit :=!\ \;%
    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t6-c}(0.01,-0.10)(-0.02,0.28)%
    \dadada%
    \tikzmarkend{t6-c}%
    \lsti!.!%
}
&
\multicolumn{4}{@{}l@{}}{
\goodTerm{
    \mathit{self} :
    \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A6-b}(0.00,-0.15)(0.00,0.30)%
    \big[\texttt{tm} : \cU,\ \texttt{tm\_unit} : \El{\texttt{tm}},\ \dadada\big]
    \tikzmarkend{A6-b}%
}{
    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t6-b}(-0.02,-0.10)(-0.02,0.28)
%   \lambda\_.\,\lambda\_.\,\lambda x.\,\lambda t.\,\mathit{self}.\texttt{tm\_unit}
    \dadada
    \tikzmarkend{t6-b}
}{
    \tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T6-b}(-0.05,-0.30)(-0.02,0.46)
%   \CaseSig{\top}{\bot}{
    \texttt{CaseTy}\left(\top,\bot,
        \begin{minipage}{.15\textwidth}%
        \vspace{-7.5pt}%
        \fontsize{8.0}{9.2}\selectfont%
        \begin{gather*}
        \begin{array}{@{}l@{}}
            \El{\mathit{self}.\texttt{tm}} \to
            \\
            \mathit{id} \to \El{\mathit{self}.\texttt{tm}}
        \end{array}
        \end{gather*}
        \end{minipage}
    \right)
%   }
    \tikzmarkend{T6-b}
}
}
\\
\cellcolor{Bkg-color-2}
\lsti!Case!\ \dadada\ \ 
\lsti!Case!\ \dadada\ \ 
\lsti!Case!\ \dadada%
&
\dadada
\\
\cellcolor{Bkg-color}
&
\lsig_{10} \coloneqq \LSigAdd{\lsig_9}{\seal_{10}}{
    \tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T10-a}(-0.02,-0.10)(-0.02,0.28)
    T_{10}
    \tikzmarkend{T10-a}
}
&
\lkg_{10} \coloneqq \LkgAdd{\lkg_9}{
    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t10-a}(-0.02,-0.10)(-0.02,0.28)
    t_{10}
    \tikzmarkend{t10-a}
}
&
\goodTerm{\cdot}{\lkg_{10}}{\TyLkg{\lsig_{10}}}
&
\goodTerm{x : \TyTkg{\lsig_9}}{\seal_{10}}{{
    \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A10-a}(-0.02,-0.10)(-0.05,0.28)%
    A_{10}
    \tikzmarkend{A10-a}
}{}}
\\
\cellcolor{Bkg-color}
\multirow{-2}{*}{%
    \lsti!End!\ \ %
    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t10-c}(0.01,-0.10)(-0.02,0.28)%
    \lsti!subst!%
    \tikzmarkend{t10-c}%
    \lsti!.!%
}
&
\multicolumn{4}{@{}l@{}}{
\goodTerm{
    \mathit{self} :
    \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A10-b}(0.00,-0.32)(0.00,0.47)%
    \!\left[{%
    \begin{minipage}{.31\textwidth}%
    \vspace{-7.5pt}%
    \fontsize{8.0}{9.2}\selectfont%
    \begin{gather*}
    \begin{array}{@{}l@{}}
        \texttt{tm} : \TyS{\wcode{\wsig_{\texttt{tm}}}},\ 
        \texttt{tm\_unit} : \El{\texttt{tm}},\ 
        \dadada,
        \smallskip\\
        \texttt{subst\_tm\_unit} : \CaseSig{\top}{\bot}{\dadada},\ 
        \dadada
    \end{array}
    \end{gather*}
    \end{minipage}
    }\right]\!
    \tikzmarkend{A10-b}%
}{
    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t10-b}(-0.02,-0.10)(-0.02,0.28)
    \lambda t.\,\Wrec{\wsig_{\texttt{tm}}}{\dadada}{t}
    \tikzmarkend{t10-b}
}{
    \begin{minipage}{.15\textwidth}%
    \vspace{-7.5pt}%
    \fontsize{8.0}{9.2}\selectfont%
    \begin{gather*}
    \tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T10-b}(-0.02,-0.45)(-0.02,0.60)
    \begin{array}{@{}l@{}}
        \El{\mathit{self}.\texttt{tm}} \to
        \\
        \El{\mathit{self}.\texttt{tm}} \to
        \\
        \mathit{id} \to \El{\mathit{self}.\texttt{tm}}
    \end{array}
    \tikzmarkend{T10-b}
    \end{gather*}
    \end{minipage}
}
}
\\
\cellcolor{Bkg-color-2}
\dadada
&
\dadada
\\
\cellcolor{Bkg-color}
\lsti!End !%
\tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t-STLC-c}(-0.02,-0.10)(-0.02,0.28)%
\lsti!STLC!%
\tikzmarkend{t-STLC-c}%
\lsti!.!
&
\goodTerm{\cdot}{
    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t-STLC-b}(-0.02,-0.10)(-0.02,0.28)%
    \Tkg{\lkg_{53}}
    \tikzmarkend{t-STLC-b}
}{\TyTkg{\lsig_{53}}}
\end{tabular}

\caption{\TT encoding of the \texttt{STLC} family from \cref{fig:stlc-mechanized}.
Each row $\goodTerm{\mathit{self}\!:\!
\tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A-i-cap}(-0.02,-0.10)(-0.02,0.28)
A_i%
\tikzmarkend{A-i-cap}%
}{%
\tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t-i-cap}(-0.02,-0.10)(-0.02,0.28)%
t_i%
\tikzmarkend{t-i-cap}%
}{%
\tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T-i-cap}(-0.02,-0.10)(-0.02,0.28)%
T_i%
\tikzmarkend{T-i-cap}%
}$ types a field.
The type $A_i$ controls how the typing of field $t_i$ sees the types of the fields prior to $t_i$.
}
\label{fig:stlc-linkage-typing}
\end{figure}