### 2023-03-21 conclusion.tex
tweaks

It is hard to write modular, extensible code and proofs.
We have presented a solution that equips a proof assistant with
linguistic facilities for family polymorphism.
The language design ensures that the expressive power brought by family polymorphism
is in harmony with the strictness of a proof assistant, while incurring
low cognitive overhead and allowing an idiomatic programming experience.
We implement the design via a translation to Coq and demonstrate its applicability
using case studies.
A novel dependent type theory formalizes the essence of the language mechanism
and is shown to enjoy consistency and canonicity.
Future work can explore ways to
further improve the practicality and expressivity of the language mechanism.
%Generalizing the mechanism to support nested family polymorphism
%will make extensibility scale to sizable proof engineering efforts and
%thus make exciting directions for future work.