### 2023-03-23 stlc-linkage-typing-named.tex
revamp figure 8

\begin{figure}
\small
\renewcommand*{\arraystretch}{1.15}
\setlength\tabcolsep{0pt}
\newcolumntype{L}{>{$}l<{$}}
\newcolumntype{C}{>{$}c<{$}}
\newcolumntype{R}{>{$}r<{$}}
%
\colorlet{Bkg-color}{black!8}
\colorlet{Bkg-color-2}{white}
%

\begingroup
\renewcommand*{\arraystretch}{1.55}

\fcolorbox{black}{faint-gray}{
\begin{minipage}{25ex}%
\vspace{-1.2ex}%
\begin{gather*}%
\begin{array}{@{}l@{}}
\goodTerm{x :
    \tikzmarkin[disable rounded corners=true,set fill color=P-color,set border color=P-color]{Pi-c}(0.00,-0.15)(0.00,0.30)%
    \TyTkg{\lsig_{i-1}}
    \tikzmarkend{Pi-c}
}{\seal_i}{
    \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{Ai-c}(0.00,-0.15)(0.00,0.30)%
    A_i
    \tikzmarkend{Ai-c}
}
\\
\goodTerm{\mathit{self} :
    \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{Ai-b}(0.00,-0.15)(0.00,0.30)%
    A_i
    \tikzmarkend{Ai-b}
}{
    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{ti-b}(-0.02,-0.10)(-0.02,0.28)
    t_i
    \tikzmarkend{ti-b}
}{
    \tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{Ti-b}(0.00,-0.15)(0.00,0.30)
    T_i
    \tikzmarkend{Ti-b}
}
\\
\text{\scriptsize shown below for $i\in\{1,2,6,10\}$}
\end{array} 
\end{gather*}
\end{minipage}
}%
\hfill
\fcolorbox{black}{faint-gray}{
\begin{minipage}{25ex}%
\vspace{-1.2ex}%
\begin{gather*}%
\begin{array}{@{}l@{}}
\lsig_i \coloneq \LSigAdd{\lsig_{i-1}}{x.\seal_i}{\mathit{self}.\,
    \tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{Ti-a}(0.00,-0.15)(0.00,0.30)
    T_i
    \tikzmarkend{Ti-a}
}
\\
\lkg_i \coloneq \LkgAdd{\lkg_{i-1}}{\mathit{self}.\,
    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{ti-a}(-0.02,-0.10)(-0.02,0.28)
    t_i
    \tikzmarkend{ti-a}
}
\\
\goodTerm{\cdot}{\lkg_i}{\TyLkg{\lsig_i}}
\end{array} 
\end{gather*}
\end{minipage}
}%
\hfill
\fcolorbox{black}{faint-gray}{
\begin{minipage}{38.5ex}%
\vspace{-1.2ex}%
\begin{gather*}%
\begin{array}{@{}l@{}}
\TyTkg{\lsig_i} \coloneq \nTySigma{x}{\TyTkg{\lsig_{i-1}}}{\nsub{T_i}{\mathit{self}}{\seal_i}}
\\
\Tkg{\lkg_i} \coloneq \big(\Tkg{\lkg_{i-1}},
    \nsub{t_i}{\mathit{self}}{\nsub{\seal_i}{x}{\Tkg{\lkg_{i-1}}}}\!
\big)
\\
\goodTerm{\cdot}{\Tkg{\lkg_i}}{\TyTkg{\lsig_i}}
\end{array} 
\end{gather*}
\end{minipage}
}
\endgroup

\vspace{2ex}

\begin{tabular}{@{}l@{\ \ \ }L@{\quad}L@{\quad}L@{\quad}L@{}}
\cellcolor{Bkg-color}
\lsti!Family STLC.!
&
\lsig_0 \coloneqq \LSigEmp
&
\lkg_0 \coloneqq \LkgEmp
%&
%\goodTerm{\cdot}{\lkg_0}{\TyLkg{\lsig_0}}
&
\\
\\[-8pt]
\cellcolor{Bkg-color}
&
\multicolumn{4}{@{}l@{}}{
\goodTerm{
    \mathit{x} : 
    \tikzmarkin[disable rounded corners=true,set fill color=P-color,set border color=P-color]{P0-c}(0.00,-0.15)(0.00,0.30)%
    \big[\big]
    \tikzmarkend{P0-c}%
}{
    \seal_1
}{
    \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A1-c}(0.00,-0.15)(0.00,0.30)%
    \big[\big]
    \tikzmarkend{A1-c}%
} 
}
\\[2pt]
\cellcolor{Bkg-color}
&
\multicolumn{4}{@{}l@{}}{
\goodTerm{
    \mathit{self} : 
    \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A1-b}(0.00,-0.15)(0.00,0.30)%
    \big[\big]
    \tikzmarkend{A1-b}%
}{
    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t1-b}(-0.02,-0.10)(-0.02,0.28)
    \wcode{\wsig_{\texttt{tm}}}
    \tikzmarkend{t1-b}
}{
    \tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T1-b}(-0.02,-0.10)(-0.02,0.28)
    \TyS{\wcode{\wsig_{\texttt{tm}}}}
    \tikzmarkend{T1-b}
} 
}
\\[2pt]
\cellcolor{Bkg-color}
\multirow{-3}{*}{%
\lsti!FInductive!\ \ 
    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t1-c}(0.03,-0.10)(-0.02,0.28)%
    \lsti!tm!%
    \tikzmarkend{t1-c}%
\ \;\lsti!:=!}
&
\lsig_1 \coloneqq \LSigAdd{\lsig_0}{x.\seal_1}{
    \mathit{self}.\,
    \tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T1-a}(-0.02,-0.10)(-0.02,0.28)
    T_1
    \tikzmarkend{T1-a}
}
&
\lkg_1 \coloneqq \LkgAdd{\lkg_0}{
    \mathit{self}.\,
    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t1-a}(-0.02,-0.10)(-0.02,0.28)
    t_1
    \tikzmarkend{t1-a}
}
%&
%\goodTerm{\cdot}{\lkg_1}{\TyLkg{\lsig_1}}
%&
%\goodTerm{x : \TyTkg{\lsig_0}}{\seal_1}{{
%    \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A1-a}(-0.02,-0.10)(-0.05,0.28)
%    A_1
%    \tikzmarkend{A1-a}
%}{}}
\\
\\[-8pt]
\cellcolor{Bkg-color}
&
\multicolumn{4}{@{}l@{}}{
\goodTerm{
    \mathit{x} :
    \tikzmarkin[disable rounded corners=true,set fill color=P-color,set border color=P-color]{P2-c}(0.00,-0.15)(0.00,0.30)%
    \big[\texttt{tm} : \TyS{\wcode{\wsig_{\texttt{tm}}}}\big]
    \tikzmarkend{P2-c}%
}{
    \seal_2
}{
    \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A2-c}(0.00,-0.15)(0.00,0.30)
    \big[\texttt{tm} : \TyS{\wcode{\wsig_{\texttt{tm}}}}\big]
    \tikzmarkend{A2-c}
}
}
\\[2pt]
\cellcolor{Bkg-color}
&
\multicolumn{4}{@{}l@{}}{
\goodTerm{
    \mathit{self} :
    \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A2-b}(0.00,-0.15)(0.00,0.30)%
    \big[\texttt{tm} : \TyS{\wcode{\wsig_{\texttt{tm}}}}\big]
    \tikzmarkend{A2-b}%
}{
    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t2-b}(-0.02,-0.10)(-0.02,0.28)
    \Wsup{}{\wsig_{\texttt{tm}}}{\top}{x.\bot}
    \tikzmarkend{t2-b}
}{
    \tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T2-b}(-0.02,-0.10)(-0.02,0.28)
    \El{\getfield{\mathit{self}}{\texttt{tm}}}
    \tikzmarkend{T2-b}
}
}
\\[2pt]
\cellcolor{Bkg-color}
\multirow{-3}{*}{%
    \lsti!|!\ %
    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t2-c}(0.01,-0.10)(-0.02,0.28)%
    \lsti!tm_unit!%
    \tikzmarkend{t2-c}%
    \ \ \lsti!: tm!
}
&
\lsig_2 \coloneqq \LSigAdd{\lsig_1}{x.\seal_2}{
    \mathit{self}.\,
    \tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T2-a}(-0.02,-0.10)(-0.02,0.28)
    T_2
    \tikzmarkend{T2-a}
}
&
\lkg_2 \coloneqq \LkgAdd{\lkg_1}{
    \mathit{self}.\,
    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t2-a}(-0.02,-0.10)(-0.02,0.28)
    t_2
    \tikzmarkend{t2-a}
}
%&
%\goodTerm{\cdot}{\lkg_2}{\TyLkg{\lsig_2}}
%&
%\goodTerm{x : \TyTkg{\lsig_1}}{\seal_2}{{
%    \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A2-a}(-0.02,-0.10)(-0.05,0.28)
%    A_2
%    \tikzmarkend{A2-a}
%}{}}
\\[2pt]
\cellcolor{Bkg-color-2}
\lsti!| !\dadada\ \ \lsti!| !\dadada\ \ \lsti!| !\dadada\lsti!.!
&
\dadada
\\[4pt]
\cellcolor{Bkg-color-2}
\lsti!FRecursion subst!\ \;\dadada\lsti!.!
&
\multicolumn{4}{@{}l@{}}{
\textit{\color{codecomment-color}%
    $\wcode{\wsig_{\texttt{tm}}}$ is abstracted into $\cU$ in the typing
    context (i.e., $A_6$) of the case handler below:%
}
}
\\
\cellcolor{Bkg-color}
&
\multicolumn{4}{@{}l@{}}{
\goodTerm{
    \mathit{x} :
    \tikzmarkin[disable rounded corners=true,set fill color=P-color,set border color=P-color]{P6-c}(0.00,-0.15)(0.00,0.30)%
    \big[\texttt{tm} : \TyS{\wcode{\wsig_{\texttt{tm}}}},\ \texttt{tm\_unit} : \El{\texttt{tm}},\ \dadada\big]
    \tikzmarkend{P6-c}%
}{
    \seal_6
}{
    \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A6-c}(0.01,-0.10)(-0.02,0.28)
    \big[\texttt{tm} : \cU,\ \texttt{tm\_unit} : \El{\texttt{tm}},\ \dadada\big]
    \tikzmarkend{A6-c}
}
}
\\[2pt]
\cellcolor{Bkg-color}
&
\multicolumn{4}{@{}l@{}}{
\goodTerm{
    \mathit{self} :
    \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A6-b}(0.00,-0.15)(0.00,0.30)%
    \big[\texttt{tm} : \cU,\ \texttt{tm\_unit} : \El{\texttt{tm}},\ \dadada\big]
    \tikzmarkend{A6-b}%
}{
    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t6-b}(-0.02,-0.10)(-0.02,0.28)
%   \lambda\_.\,\lambda\_.\,\lambda x.\,\lambda t.\,\mathit{self}.\texttt{tm\_unit}
    \dadada
    \tikzmarkend{t6-b}
}{
    \tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T6-b}(-0.05,-0.30)(-0.02,0.46)
%   \CaseSig{\top}{\bot}{
    \texttt{CaseTy}\left(\top,\bot,
        \begin{minipage}{.15\textwidth}%
        \vspace{-7.5pt}%
        \fontsize{8.0}{9.2}\selectfont%
        \begin{gather*}
        \begin{array}{@{}l@{}}
            \El{\getfield{\mathit{self}}{\texttt{tm}}} \to
            \\
            \mathit{id} \to \El{\getfield{\mathit{self}}{\texttt{tm}}}
        \end{array}
        \end{gather*}
        \end{minipage}
    \right)
%   }
    \tikzmarkend{T6-b}
}
}
\\[2pt]
\cellcolor{Bkg-color}
\multirow{-4}{*}{%
    \lsti!Case tm_unit :=!\ \;%
    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t6-c}(0.01,-0.10)(-0.02,0.28)%
    \dadada%
    \tikzmarkend{t6-c}%
    \lsti!.!%
}
&
\lsig_6 \coloneqq \LSigAdd{\lsig_5}{x.\seal_6}{
    \mathit{self}.\,
    \tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T6-a}(-0.02,-0.10)(-0.02,0.28)
    T_6
    \tikzmarkend{T6-a}
}
&
\lkg_6 \coloneqq \LkgAdd{\lkg_5}{
    \mathit{self}.\,
    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t6-a}(-0.02,-0.10)(-0.02,0.25)
    t_6
    \tikzmarkend{t6-a}
}
%&
%\goodTerm{\cdot}{\lkg_6}{\TyLkg{\lsig_6}}
%&
%\goodTerm{x : \TyTkg{\lsig_5}}{\seal_6}{{
%    \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A6-a}(-0.02,-0.10)(-0.05,0.28)
%    A_6
%    \tikzmarkend{A6-a}
%}}
\\[2pt]
\cellcolor{Bkg-color-2}
\lsti!Case!\ \dadada\ \ 
\lsti!Case!\ \dadada\ \ 
\lsti!Case!\ \dadada%
&
\dadada
\\
\cellcolor{Bkg-color}
&
\multicolumn{4}{@{}l@{}}{
\goodTerm{
    \mathit{x} :
    \tikzmarkin[disable rounded corners=true,set fill color=P-color,set border color=P-color]{P10-c}(0.00,-0.32)(0.00,0.47)%
    \!\left[{%
    \begin{minipage}{.31\textwidth}%
    \vspace{-7.5pt}%
    \fontsize{8.0}{9.2}\selectfont%
    \begin{gather*}
    \begin{array}{@{}l@{}}
        \texttt{tm} : \TyS{\wcode{\wsig_{\texttt{tm}}}},\ 
        \texttt{tm\_unit} : \El{\texttt{tm}},\ 
        \dadada,
        \smallskip\\
        \texttt{subst\_tm\_unit} : \CaseSig{\top}{\bot}{\dadada},\ 
        \dadada
    \end{array}
    \end{gather*}
    \end{minipage}
    }\right]\!
    \tikzmarkend{P10-c}%
}{
    \seal_{10}
}{
    \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A10-c}(0.00,-0.32)(0.00,0.47)%
    \!\left[{%
    \begin{minipage}{.31\textwidth}%
    \vspace{-7.5pt}%
    \fontsize{8.0}{9.2}\selectfont%
    \begin{gather*}
    \begin{array}{@{}l@{}}
        \texttt{tm} : \TyS{\wcode{\wsig_{\texttt{tm}}}},\ 
        \texttt{tm\_unit} : \El{\texttt{tm}},\ 
        \dadada,
        \smallskip\\
        \texttt{subst\_tm\_unit} : \CaseSig{\top}{\bot}{\dadada},\ 
        \dadada
    \end{array}
    \end{gather*}
    \end{minipage}
    }\right]\!
    \tikzmarkend{A10-c}%
}
}
\\[2pt]
\cellcolor{Bkg-color}
&
\multicolumn{4}{@{}l@{}}{
\goodTerm{
    \mathit{self} :
    \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A10-b}(0.00,-0.32)(0.00,0.47)%
    \!\left[{%
    \begin{minipage}{.31\textwidth}%
    \vspace{-7.5pt}%
    \fontsize{8.0}{9.2}\selectfont%
    \begin{gather*}
    \begin{array}{@{}l@{}}
        \texttt{tm} : \TyS{\wcode{\wsig_{\texttt{tm}}}},\ 
        \texttt{tm\_unit} : \El{\texttt{tm}},\ 
        \dadada,
        \smallskip\\
        \texttt{subst\_tm\_unit} : \CaseSig{\top}{\bot}{\dadada},\ 
        \dadada
    \end{array}
    \end{gather*}
    \end{minipage}
    }\right]\!
    \tikzmarkend{A10-b}%
}{
    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t10-b}(-0.02,-0.10)(-0.02,0.28)
    \lambda x.\,\Wrec{\wsig_{\texttt{tm}}}{\dadada}{x}
    \tikzmarkend{t10-b}
}{
    \begin{minipage}{.15\textwidth}%
    \vspace{-7.5pt}%
    \fontsize{8.0}{9.2}\selectfont%
    \begin{gather*}
    \tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T10-b}(-0.02,-0.45)(-0.02,0.57)
    \begin{array}{@{}l@{}}
        \El{\getfield{\mathit{self}}{\texttt{tm}}} \to
        \\
        \El{\getfield{\mathit{self}}{\texttt{tm}}} \to
        \\
        \mathit{id} \to \El{\getfield{\mathit{self}}{\texttt{tm}}}
    \end{array}
    \tikzmarkend{T10-b}
    \end{gather*}
    \end{minipage}
}
}
\\[2pt]
\cellcolor{Bkg-color}
\multirow{-5}{*}{%
    \lsti!End!\ \ %
    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t10-c}(0.01,-0.10)(-0.02,0.28)%
    \lsti!subst!%
    \tikzmarkend{t10-c}%
    \lsti!.!%
}
&
\lsig_{10} \coloneqq \LSigAdd{\lsig_9}{x.\seal_{10}}{
    \mathit{self}.\,
    \tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T10-a}(-0.02,-0.10)(-0.02,0.28)
    T_{10}
    \tikzmarkend{T10-a}
}
&
\lkg_{10} \coloneqq \LkgAdd{\lkg_9}{
    \mathit{self}.\,
    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t10-a}(-0.02,-0.10)(-0.02,0.28)
    t_{10}
    \tikzmarkend{t10-a}
}
%&
%\goodTerm{\cdot}{\lkg_{10}}{\TyLkg{\lsig_{10}}}
%&
%\goodTerm{x : \TyTkg{\lsig_9}}{\seal_{10}}{{
%    \tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A10-a}(-0.02,-0.10)(-0.05,0.28)%
%    A_{10}
%    \tikzmarkend{A10-a}
%}{}}
\\[2pt]
\cellcolor{Bkg-color-2}
\dadada
&
\dadada
\\
\cellcolor{Bkg-color}
\lsti!End !%
\tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t-STLC-c}(-0.02,-0.10)(-0.02,0.28)%
\lsti!STLC!%
\tikzmarkend{t-STLC-c}%
\lsti!.!
&
\goodTerm{\cdot}{
    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t-STLC-b}(-0.02,-0.10)(-0.02,0.28)%
    \Tkg{\lkg_{53}}
    \tikzmarkend{t-STLC-b}
}{\TyTkg{\lsig_{53}}}
\end{tabular}

%\begin{tabular}{ll}
%  \hline
%  $\TyTkg{\lsig_1} = \nTySigma{\_}{\top}{\TyS{\wcode{\wsig_{\texttt{tm}}}}} $ &  
%  $\TyTkg{\lsig_2} = \nTySigma{\_}{\TyTkg{\lsig_1}}{\El{\wcode{\wsig_{\texttt{tm}}}}}$ \\  
%  $\TyTkg{\lsig_6} = \nTySigma{P_5}{\TyTkg{\lsig_5}}{\CaseSig{\top}{\bot}{\El{\getfield{P_5}{"tm"}} \to \mathit{id} \to \El{\getfield{P_5}{"tm"}}}}$ & \\ 
%  $\TyTkg{\lsig_{10}} = \nTySigma{\_}{\TyTkg{\lsig_9}}{\El{\wcode{\wsig_{\texttt{tm}}}} \to \El{\wcode{\wsig_{\texttt{tm}}}} \to \mathit{id} \to \El{\wcode{\wsig_{\texttt{tm}}}}}$ & \\ 
%\end{tabular}

\caption{\TT encoding of the \texttt{STLC} family from \cref{fig:stlc-mechanized}.
Each row $\goodTerm{\mathit{self}\!:\!
\tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A-i-cap}(-0.02,-0.10)(-0.02,0.28)
A_i%
\tikzmarkend{A-i-cap}%
}{%
\tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t-i-cap}(-0.02,-0.10)(-0.02,0.28)%
t_i%
\tikzmarkend{t-i-cap}%
}{%
\tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T-i-cap}(-0.02,-0.10)(-0.02,0.28)%
T_i%
\tikzmarkend{T-i-cap}%
}$ types a field.
The type $A_i$ controls how the typing of field $t_i$ sees the types of the fields prior to $t_i$.
}
\label{fig:stlc-linkage-typing}

\end{figure}