### 2023-03-31 body.tex
Set out the frame

\input{frontmatters.tex}

\newpage
\pagenumbering{arabic}
\setcounter{page}{1}
\chapter{Family Polymorphism as Coq Plugin}

\ifreport
\thispagestyle{tr-style}
\fi

% lstlisting coq style (inspired from a file of Assia Mahboubi)
% https://tex.stackexchange.com/questions/434523/coq-syntax-highlighting
\section{Introduction}
\label{sec:intro}
\input{introduction.tex}

% Concept 
% Contribution

\section{Design Requirements and Challenges}
% what feature we want
% what problem/formalization we want to achieve
% Before/After Comparison of examples (with/without Fampoly)
\label{sec:background+challenge}
\input{challenge.tex}

\section{Language Design}
% we need to make this section before Metatheory
% because a lot of our formulation in the metatheory
%   is inspired by the implementation
\label{sec:lang-design}
\input{lang-design.tex}

\section{Compiling Family Polymorphism to Parameterized Modules}
% we need to make this section before Metatheory
% because a lot of our formulation in the metatheory
%   is inspired by the implementation
\label{sec:coqimpl}
\input{coqimpl.tex}

\section{A Preview of \TT: A Core Dependent Type Theory}
\label{sec:metatheory2}
\input{metatheory-condensed.tex}

\section{Case Studies}
\label{sec:casestudies}
\input{casestudies.tex}

\section{Related Work}\label{sec:related-work}
\input{related-work.tex}

\section{Conclusion}
\label{sec:conclusion}

\input{conclusion.tex}

\newpage

\chapter{Technicality of \TT}

\section{Syntax and Semantic Models of \TT}
\label{sec:complete-fmltt}
% \input{all-typing-rule.tex} -- we make all-typing rules into metatheory2.tex
\input{metatheory2.tex}

% \newpage

\section{Using \TT's Linkage Transformers to Model a Derived Family}
\label{sec:lt-in-action}
\input{stlc-linkTran-typing.tex}

% \newpage

% \section{Definitive Version of the Formalization and Proofs}
% \label{sec:agda-files}
% \input{agda-files.tex}

%\section{Metatheory, in detail}
%\label{sec:metatheory}
%\input{metatheory.tex}
%
%\section{Example in Metatheory}
%\label{sec:coqexample}
%\input{coq-example.tex}

% \fi

% \ifacknowledgments
% \section*{Acknowledgments}
% We thank the reviewers for their many useful comments.  We thank
% Anastasiya Kravchuk-Kirilyuk, Yao Li, Andrew Myers, and Zhixuan Yang
% for valuable discussions and feedback.
% This work was supported in part by an Amazon research award on automated
% reasoning.
% The views and opinions expressed are those of the authors and do not necessarily
% reflect the position of any funding agency.
% \fi
\input{backmatters.tex}