### 2023-03-31 frontmatters.tex
Add Front Matter

% Title Page

\vspace*{\fill}
\begin{center}
  \textsc{\Huge \thetitle}
\end{center}

\vspace*{\fill}

\newpage

% Author's Declaration
\vspace*{\fill}

\chapterN{Author's Declaration}

\vspace*{\fill}

\newpage

% Statement of contributions

\vspace*{\fill}

\chapterN{Statement of contributions}

\vspace*{\fill}

\newpage

% Abstract
\vspace*{\fill}
\chapterN{Abstract}

\input{abstract}
\vspace*{\fill}

\newpage

% Acknowledgements

\vspace*{\fill}
\chapterN{Acknowledgments}

We thank Anastasiya Kravchuk-Kirilyuk, Yao Li, Andrew Myers, and Zhixuan Yang
for valuable discussions and feedback.
This work was supported in part by an Amazon research award on automated
reasoning.
The views and opinions expressed are those of the authors and do not necessarily
reflect the position of any funding agency.
\vspace*{\fill}

\newpage
\tableofcontents