### 2023-04-01 frontmatters.tex
Fix.

% Title Page
\pagenumbering{roman}
\vspace*{\fill}
\begin{center}
  \textsc{\Huge \thetitle}
\end{center}

\vspace*{\fill}

\newpage

\setcounter{page}{1}

% Author's Declaration
\vspace*{\fill}

\chapterN{Author's Declaration}

This thesis consists of material all of which I authored or co-authored: see Statement of Contributions included in the thesis. This is a true copy of the thesis, including any required final revisions, as accepted by my examiners.

\vspace*{\fill}

\newpage

% Statement of contributions

\vspace*{\fill}

\chapterN{Statement of contributions}

Ende Jin and Prof. Yizhou Zhang co-authored Chapter 1.

Ende Jin is the sole author of Chapter 2, which were written under the  under the supervision of Prof. Yizhou Zhang.

Both Chapter 1 and Chapter 2 are already published result.

\noindentparagraph{Research presented in Chapter 1 and 2:} Ende Jin is the primary investigator of both theoretical development and practical implementation of the prototype. This work was supported in part by an Amazon research award on automated
reasoning.

\vspace*{\fill}

\newpage

% Abstract
\vspace*{\fill}
\chapterN{Abstract}

\input{abstract}
\vspace*{\fill}

\newpage

% Acknowledgements

\vspace*{\fill}
\chapterN{Acknowledgments}

We thank Anastasiya Kravchuk-Kirilyuk, Yao Li, Andrew Myers, and Zhixuan Yang
for valuable discussions and feedback.
This work was supported in part by an Amazon research award on automated
reasoning.
The views and opinions expressed are those of the authors and do not necessarily
reflect the position of any funding agency.
\vspace*{\fill}

\newpage
\tableofcontents