### 2023-04-01 title.tex
Adding List of Figures.

\def\thetitle{%
Application and Theory of \mbox{Family Polymorphism} in \mbox{Proof Assistant}
% Extensible Meta\-theory Mechanization via \mbox{Family Polymorphism}%
%Family Polymorphism for Extensible Metatheory Mechanization in a Type Theory with Inductive Types
}

\def\thesubtitle{%
Technical Report with Appendices
}

\ifreport
\title{\textls[-0]{\thetitle}}
\subtitle{\thesubtitle}
\else
\title{\textls[-0]{\thetitle}}
\fi

%\subtitle{\thesubtitle}

\newcommand{\TT}{\mbox{\textls[-20]{FMLTT}}\xspace}
\newcommand{\Lang}{\mbox{\scshape\textls[-10]{fpop}}\xspace}