### 2023-04-02 title.tex
TR

\def\thetitle{%
Application and Theory of \mbox{Family Polymorphism} in \mbox{Proof Assistant}
% Extensible Meta\-theory Mechanization via \mbox{Family Polymorphism}%
%Family Polymorphism for Extensible Metatheory Mechanization in a Type Theory with Inductive Types
}

\ifreport
\title{\thetitle: Technical Report}
\else
\title{\thetitle}
\fi

\fancypagestyle{tr-style}{
    \fancyhf{}
    \fancyhead[L]{\small\sffamily Technical Report CS-2023-01, School of Computer Science, University of Waterloo}
}

\newcommand{\TT}{\mbox{\textls[-20]{FMLTT}}\xspace}
\newcommand{\Lang}{\mbox{\scshape\textls[-10]{fpop}}\xspace}