### 2023-06-08 frontmatters.tex
Fix frontmatters.

% Title Page
\pagenumbering{roman}

\input{titlepage.tex}

\newpage

\setcounter{page}{1}

% Author's Declaration
\vspace*{\fill}

\chapterN{Author's Declaration}

This thesis consists of material all of which I authored or co-authored: see Statement of Contributions included in the thesis. This is a true copy of the thesis, including any required final revisions, as accepted by my examiners.

I understand that my thesis may be made electronically available to the public.

\vspace*{\fill}

\newpage

% Statement of contributions

\vspace*{\fill}

\chapterN{Statement of contributions}

Ende Jin was the sole author of \cref{sec:complete-fmltt}, which were written under the supervision of Prof. Yizhou Zhang.

Jin E, Amin N, Zhang Y. Extensible Metatheory Mechanization via Family Polymorphism: Technical Report. Technical Report CS-2023-01. School of Computer Science, University of Waterloo. https://cs. uwaterloo. ca/sites/ca. computer-science/files/uploads/files/cs-2023-01. pdf; 2023. \\

This thesis consists in part of a manuscript written for publication. The exception to sole authorship of material is as follows:

\noindentparagraph{Research presented in rest of the thesis:}
The research was conducted at the University of Waterloo by Ende Jin under the supervision of Dr. Yizhou Zhang. Ende Jin proposed the primary design of the language, prototyped the Coq plugin for realizing the language, implemented and evaluated the case studies, and formalized and proved meta-theoretic arguments for the core type theory. Ende Jin estabilished the complete language design with consultations from Dr. Yizhou Zhang.

Jin E, Amin N, Zhang Y. Extensible Metatheory Mechanization via Family Polymorphism. Proc. of the ACM on Programming Languages (PACMPL). 2023;7. DOI: https://doi.org/10.1145/3591286

\vspace*{\fill}

\newpage

% Abstract
\vspace*{\fill}
\chapterN{Abstract}

\input{abstract}
\vspace*{\fill}

\newpage

% Acknowledgements

\vspace*{\fill}
\chapterN{Acknowledgments}

I am grateful to the reviewers for their many useful comments.  I thank
Benjamin Delaware,
Anastasiya Kravchuk-Kirilyuk, Yao Li, Andrew Myers, Zhixuan Yang, Prabhakar Ragde, Werner Dietl
for valuable discussions and feedback. I also thank my advisors: Yizhou Zhang and Ondřej Lhoták for their patience and guidance.
This work was supported in part by an Amazon research award on automated
reasoning.
The views and opinions expressed are those of the author and do not necessarily
reflect the position of any funding agency.

\vspace*{\fill}

\newpage
\tableofcontents

\newpage

\vspace*{\fill}

\listoffigures

\vspace*{\fill}

\newpage

\vspace*{\fill}
\chapterN{List of Abbreviations}

\input{listofabbrev.tex}

\vspace*{\fill}

% we omit list of symbols first
% \newpage

% \vspace*{\fill}
% \chapterN{List of Symbols}

\vspace*{\fill}